---
title: "Data 607 - Week 3 Assignment"
author: "Jeff Parks"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

### 1. Provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r setup, message=FALSE}
library(tidyverse)
library(fivethirtyeight)

data("college_all_ages")
df <- college_all_ages
str_subset((df$major),regex("DATA|STATISTICS", ignore_case=TRUE))
```

---

### 2. Write code that transforms the data below:

> [1] "bell pepper"  "bilberry"     "blackberry"   "blood orange" 
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"

Into a format like this:
  
> c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r, message=FALSE, warning=FALSE}
berries <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

s <- str_extract_all(str_squish(berries),'(?<=")([:alpha:]|[:space:]){2,}(?=")')
u <- str_c(s, sep='"')
writeLines(u)
```

---

### 3. Describe, in words, what these expressions will match:

#### (.)\1\1 
Will match nothing .. no required quotes to define the regex pattern, and single backslash only for escape characters.

#### "(.)(.)\\\2\\\1" 
Will match the first case where two single characters appear in order, and then appear immediately again in reverse order.
```{r}
s <- "abba" # example
str_view(s, "(.)(.)\\2\\1")
```

#### (..)\1 
Will match nothing .. no required quotes to define the regex pattern, and single backslash only for escape characters.

#### "(.).\\\1.\\\1"
Will match the first case where a single character appears three times with any other character (except a newline) in between.
```{r}
s <- "abacab" # example
str_view(s, "(.).\\1.\\1")
```

#### "(.)(.)(.).*\\\3\\\2\\\1"
Will match the first case where a sequence of three characters appear again in reverse order, with any single character (except a newline) in between.
```{r}
s <- "abcZcba" #example
str_view(s, "(.)(.)(.).*\\3\\2\\1")
```

---

### 4. Construct regular expressions to match words that:

a. Start and end with the same character.
```{r}
s <- c('goat','llama','alpaca')
str_match(s, "^(.).*\\1$")
```

b. Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
s <- c('hall','lurch','church')
str_match(s,"^.*(..).*\\1.*$")
```

c. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
```{r}
s <- c('ten','eleven','alpaca')
str_match(s, "^.*(.).*\\1.*\\1.*$")
```