---
title: "FiveThirtyEight COVID-19 Approval Tracker"
author: "Jeff Parks"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)

# load data
polls_adjusted <- read_csv('https://raw.githubusercontent.com/jefedigital/cuny-data-607/main/covid-19-polls/data/covid_approval_polls_adjusted.csv')
df <- select(polls_adjusted, !c(modeldate,multiversions,tracking,timestamp,url))

# rename variables
df <- rename(df, poll_start = startdate)
df <- rename(df, poll_end = enddate)
df <- rename(df, sample_size = samplesize)
df <- rename(df, poll_weight = weight)
df <- rename(df, poll_influence = influence)

# fix datatypes
ctypes <- c('ffcccfdfdddddd')
df <- type_convert(df, col_types=ctypes)

# fix date separately
df$poll_start <- parse_date(df$poll_start,'%m/%d/%Y')
df$poll_end <- parse_date(df$poll_end,'%m/%d/%Y')

## fix error in population ..'v' is a data error
levels(df$population)[4] <- levels(df$population)[2] # assign v to rv
levels(df$population) <- c('unregistered','registered voter','likely voter')

df <- mutate(df, registered_voter = ifelse(population == 'unregistered', FALSE, TRUE))
df <- mutate(df, likely_voter = ifelse(population == 'likely voter', TRUE, FALSE))
```

## Introduction

FiveThirtyEight publishes a [daily tracker](https://projects.fivethirtyeight.com/coronavirus-polls/) measuring public approval of the President's handling of the COVID-19 pandemic.  

Their approach aggregates all available, scientifically-valid surveys from the field, and weights the results from each pollster using a [custom rating formula](https://projects.fivethirtyeight.com/pollster-ratings/) that factors in historical accuracy, adherence to industry best practices, and reliance on landline responses.

Let's take a quick tour.

---

## Observations

This overview uses the **covid_approval_polls_adjusted** [dataset](https://github.com/fivethirtyeight/covid-19-polls) updated on August 25, 2021.  The dataset contains results from 783 distinct surveys from 54 polling firms, beginning February 2, 2020 and ending August 24, 2021.

```{r eda, include=FALSE}
# count of polls
nrow(unique(df[,c('poll_end','pollster')])) # 783

# count of pollsters
length(unique(df$pollster)) #54

# date range
range(df$poll_start) # 2020-02-02
range(df$poll_end) # 2021-08-24
```

The survey questions vary slightly among pollsters, but roughly condense to "Do you approve or disapprove of the President's handling of the coronavirus outbreak? (Yes/No)" 

Breakdowns of respondents include political party affiliation (Democratic, Republican, Independent) and voter registration status (Unregistered Adult, Registered Voter, Likely Voter)

```{r approval_spread, include=FALSE}
df_spread_approve <- select(df, c(party,poll_end,approve_adjusted))
df_spread_approve <- df_spread_approve %>% mutate(poll_week=floor_date(poll_end, unit='week')) # get week of end_date
df_spread_approve <- df_spread_approve %>% group_by(party, poll_week) %>% summarize(avg_approval=mean(approve_adjusted)) # summarize
df_spread_approve <- df_spread_approve %>% pivot_wider(names_from=party, values_from=avg_approval) 
df_spread_approve <- df_spread_approve %>% mutate(di_approve = abs(D-I),dr_approve = abs(D-R),ir_approve = abs(I-R))
```

```{r disapproval_spread, include=FALSE}
df_spread_disapprove <- select(df, c(party,poll_end,disapprove_adjusted))
df_spread_disapprove <- df_spread_disapprove %>% mutate(poll_week=floor_date(poll_end, unit='week')) # get week of end_date
df_spread_disapprove <- df_spread_disapprove %>% group_by(party, poll_week) %>% summarize(avg_disapproval=mean(disapprove_adjusted)) # summarize
df_spread_disapprove <- df_spread_disapprove %>% pivot_wider(names_from=party, values_from=avg_disapproval)
df_spread_disapprove <- df_spread_disapprove %>% mutate(di_disapprove = abs(D-I),dr_disapprove = abs(D-R),ir_disapprove = abs(I-R))
```

```{r voter_variables, include=FALSE}
df_approve_voter <- select(df, c(party,poll_end,approve_adjusted,disapprove_adjusted,registered_voter,likely_voter))
df_approve_voter <- df_approve_voter %>% mutate(poll_week=floor_date(poll_end, unit='week'))
df_approve_voter <- df_approve_voter %>% group_by(party, poll_week, registered_voter, likely_voter) %>% summarize(avg_approval = mean(approve_adjusted), avg_disapproval=mean(disapprove_adjusted))
```

This dataset informs one of the main findings in the [FiveThirtyEight Coronavirus Polls](https://projects.fivethirtyeight.com/coronavirus-polls/}) tracker, with the finding **"Approval of Biden's response varies widely by party."**

```{r approval_party, echo=FALSE, fig.align="center"}
df_approve_voter %>%
  filter(party!='all') %>% 
  ggplot(aes(x=poll_week, y=avg_approval)) + 
  geom_line(aes(color=party)) + 
  scale_y_continuous(name = "Approval", limits = c(0, 100)) + 
  scale_color_manual(values = c('D'='blue', 'R'='red', 'I'='grey')) +
  ggtitle("Approval Rating - by Party")
```

While this conclusion is not at all surprising given our political climate, some additional questions come to mind that might be worth exploring:

---

**What does the overall picture look like, without party identification?**

```{r overall, echo=FALSE, fig.show="hold", out.width="50%"}
df_approve_voter %>%
  filter(party=='all') %>% 
  ggplot(aes(x=poll_week, y=avg_approval)) + 
  geom_line(colour='green') + 
  scale_y_continuous(name = "Approval", limits = c(0, 100)) + 
  ggtitle("Approval Rating - All Reponses")

df_approve_voter %>%
  filter(party=='all') %>% 
  ggplot(aes(x=poll_week, y=avg_disapproval)) + 
  geom_line(colour='red') + 
  scale_y_continuous(name = "Dispproval", limits = c(0, 100)) + 
  ggtitle("Dispproval Rating - All Reponses")
```

In the aggregate, voters have shifted decisively in favor of the new administration's handling of the COVID-19 pandemic .. so far.

---

**Has political polarization on this issue increased or decreased between administrations, and what is its general direction now?**

```{r spread_approve, echo=FALSE, fig.show="hold", out.width="50%"}
ggplot(df_spread_approve, aes(x=poll_week, y=dr_approve)) + 
  geom_line(colour='blue') + 
  scale_y_continuous(name = "Point Spread", limits = c(0, 100)) + 
  ggtitle("Difference between D & R Approval ratings")

ggplot(df_spread_disapprove, aes(x=poll_week, y=dr_disapprove)) + 
  geom_line(colour='blue') + 
  scale_y_continuous(name = "Point Spread", limits = c(0, 100)) + 
  ggtitle("Difference between D & R Disapproval ratings")
```

By taking the difference between the average approval andd disapproval ratings of the two main parties, we can see that overall polarization around Presidential handling of COVID-19 peaked prior to the January transition, dropped significantly during the summer, but is slowly ticking back up. 

---

**Are there significant deviations from this trend between Registered and Unregistered voters?  Between Likely and Unlikely voters?**

```{r approve_voters, echo=FALSE, fig.show="hold", out.width="50%"}
df_approve_voter %>%
  filter(party=='all') %>% 
  ggplot(aes(x=poll_week, y=avg_approval)) + 
  geom_line(aes(color=registered_voter)) + 
  scale_y_continuous(name = "Approval", limits = c(0, 100)) + 
  ggtitle("Approval Rating - Voter Registration")

df_approve_voter %>%
  filter(party=='all') %>% 
  ggplot(aes(x=poll_week, y=avg_approval)) + 
  geom_line(aes(color=likely_voter)) + 
  scale_y_continuous(name = "Approval", limits = c(0, 100)) + 
  ggtitle("Approval Rating - Voter Likelihood")
```

Surprisingly, neither voter registration nor "likely voter" status seemed have an impact on approval ratings.

---

## Conclusion

FiveThirtyEight's approach to poll analysis is fascinating.  By acknowledging the biases inherent in different methodologies, capabilities and ideological viewpoints between pollsters, they have built a fairly robust and even-handed model for understanding voter sentiment and for predicting electoral outcomes. 